NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-06-25 Manuscript/manuscriptPS.tex
Update text

Several classical molecular dynamics (MD) simulation studies are done
to understand PS headgroups, their influence on lipid bilayer properties and
interactions with ions \cite{vernier09,boettcher11,melcrova16,??}.
However, the consesus has not been reached 
due to the difficulties to interpret the experimental data \cite{??} and
the inaccuracies in simulation models at the
headgroup region \cite{botan15,catte16,ollila16}.

%Details of negatively charged lipids and their interactions
%with ions and proteins have been experimentally analyzed with
%various spectroscopic methods \cite{borle85,macdonald87,roux90,seelig90,pan14,melcrova16,??}
%NMR \cite{seelig90,??}, scattering \cite{pan14,??} and fluorescence spectroscopy
%which are often complemented with classical MD simulations \cite{pan14,melcrova16,??}.

Headgroup and glycerol backbone C-H bond order parameters
calculated from MD simulations have been
recently used to interpret the lipid structures in NMR experiments
and to validate lipid structure and ion binding in simulations of
PC lipid bilayers \cite{botan15,catte16,ollila16,ferreira16}.
In this work we apply this approach to PS lipid headgroup
in order to elucidate the structural details and ion binding
to negatively charged lipids. The results are expected to elucidate
also PS mediated signalling events because 
glycerol backbone and headgroup structure and behaviour are similar
in model membranes and in bacteria \cite{gally81,scherer87,seelig90}.

>>>>

Several classical molecular dynamics (MD) simulation studies are done
to understand PS headgroups, their influence on lipid bilayer properties and
interactions with ions \cite{vernier09,boettcher11,pan14,melcrova16,??}.
However, the recent comparisons of PC lipid headgroup and glycerol backbone
C-H bond order parameters calculated from different simulation models
revealed that improvements in the current force fields are needed
to correctly reproduce the headgroup structure and ion binding to
lipid bilayers \cite{botan15,catte16,ollila16}. The ion binding affinity
to POPC bilayer was then improved by implicitly including the electronic
polarizability using the electronic continuum correction \cite{melcr18}.
Here, we collect the set of experimentally measured lipid headgroup and
glycerol backbone C-H bond order parameters, which can be used to
evaluate the quality of headgroup structure and the ion binding affinity
in MD simulations of lipid bilayers containing PS lipids. 
The available MD simulation models of PS are then compared against
the collected experimental data. The results pave the way for the
development of MD simulation force fields that correctly describe
PS lipid headgroup structure and its interactions with ions.
Such models are expected to be useful in elucidating the biological
role of PS and other lipid headgroups because glycerol backbone and
lipid headgroups behave similarly in model membranes and in
bacteria \cite{gally81,scherer87,seelig90}.
